%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}

I want to thank everybody who supported me emotionally and professionally in the long journey towards this thesis :)

\bigskip
\noindent
Thanks to my family for their encouragements, patience and belief in me.

\bigskip
\noindent
Thanks to all my coworkers and collaborators for providing insights, help and joy during my work.

\bigskip
\noindent
Thanks especially to my supervisors Tonio Ball and Frank Hutter for providing me with the freedom and resources to tackle ambitious research and follow my dream of improving lives by using AI in the medical domain. With your work and expertise we were able to start an intense and impactful research exploration of deep learning on EEG. Let me also thank Jost Tobias Springenberg for his willingness to share his immense deep learning expertise to successfully kickstart my research journey.

\bigskip
\noindent
Thanks to Malak for showing me warmth, joy, love and support in easy and hard times.

\bigskip
\noindent
Thanks to Dephie for showing me the marvelous world, and all her encouragements, ideas, peer coaching, warmth and support.

\bigskip
\noindent
Thanks to Manuel for all his friendship, wisdom, kindness and fun throughout the many years.

\bigskip
\noindent
Thanks to Daniel for all the long-distance telephone support.

\bigskip
\noindent
Thanks to Markus, Jenny, Mirko, Linda, Esin, Edith and all others for providing me with a sense of homecoming on any Berlin trip.

\bigskip
\noindent
Thanks to Raghu for all the sport as well as his puntastic support.

\bigskip
\noindent
Thanks to Lukas and Dan for fruitful discussions and providing a fun atmosphere as officemates.


\bigskip
\noindent
Thanks to all my friends and all who have given me warmth, calmness and supported me in any way.



\bigskip

\endgroup
